---
title: CV
name: Nola Saracnei
position: "Data Ananlyst"
email: "nolan.saraceni@icloud.com"
twitter: nolanaraceni.eth
output:
  vitae::markdowncv:
    theme: kjhealy
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Case Study Questions

These three questions will guide the marketing program

-   How do annual members and casual riders use Cyclistic bikes
    differently?
-   Why would casual riders buy Cyclistic annual memberships?
-   How can Cyclistic Use digital media to influence casual riders?

#### My Report will produce the following:

1.  A clear statement of the business task
2.  a description of all data sources used
3.  Documentation of any cleaning or manipulation of data
4.  a summary of my analysis
5.  Supporting visualizations and key findings
6.  my top 3 recommendations based on my analysis

### My personal Road Map

What is the problem you are tying to solve - I am trying to solve the
problem of converting Cyclistic casual riders to annual members.

How can my insights drive business decisions - I can depict the
important KPIs. Report visuals that could track travel and purchase
behavior. Possibly incorporate incentives for when you ride certain
distances. Comparing and contrasting data in regards to riding time and
location of stations. How one might choose one station over another.
Quality of neighborhood could have an impact. (Think in and out of this
box. could possibly use outside data from other sources)

### Business Task Statement

The business task is to design marketing strategies aimed at converting
casual riders into annual members for Cyclistic, a bike-sharing program
in Chicago. The marketing analyst team needs to understand the
differences between annual members and casual riders, determine why
casual riders would purchase a membership, and explore how digital media
could impact their marketing tactics. To achieve this, they plan to
analyze the historical bike trip data from Cyclistic to identify trends
and insights.

```{r adding columns (lists day, date, month, year)}

tripdata_df$date <- as.Date(tripdata_df$start_time)
tripdata_df$month <- format(as.Date(tripdata_df$date),"%m")
tripdata_df$day <- format(as.Date(tripdata_df$date),"%d")
tripdata_df$year <- format(as.Date(tripdata_df$date),"%Y")
tripdata_df$day_of_week <- format(as.Date(tripdata_df$date),"%A")


```

```{r adding a length column in seconds}

tripdata_df$ride_length <- difftime(tripdata_df$end_time,tripdata_df$start_time)

```

```{r convert ride_length to numeric for clcuations}

tripdata_df$ride_length <- as.numeric(as.character(tripdata_df$ride_length))
is.numeric(tripdata_df$ride_length)

```

```{r Roving bad data: taking of data that contains negative second this refers to quality checking}

tripdata_df_v2 <- tripdata_df[(tripdata_df$from_station_name == 'HQ QR'| tripdata_df$ride_length>0),]

View(tripdata_df_v2)

```

```{r decriptive analysis}

mean(tripdata_df_v2$ride_length)
median(tripdata_df_v2$ride_length)
max(tripdata_df_v2$ride_length)
min(tripdata_df_v2$ride_length)

summary(tripdata_df_v2$ride_length)

```

```{r comparing casual vs members}

tripdata_df_v2$ride_length <- difftime(tripdata_df_v2$end_time,tripdata_df_v2$start_time)


aggregate(tripdata_df_v2$ride_length ~ tripdata_df_v2$usertype, FUN = mean)
aggregate(tripdata_df_v2$ride_length ~ tripdata_df_v2$usertype, FUN = max)
aggregate(tripdata_df_v2$ride_length ~ tripdata_df_v2$usertype, FUN = min)
aggregate(tripdata_df_v2$ride_length ~ tripdata_df_v2$usertype, FUN = median)


```

```{r Data aggregation}

aggregate(tripdata_df_v2$ride_length~tripdata_df_v2$usertype + tripdata_df_v2$day_of_week, FUN = mean)

```

```{r Analyzing ride data}

tripdata_df_v2%>%
  mutate(weekday = wday(start_time, label=TRUE))%>%
  group_by(usertype, weekday)%>%
  summarise(number_of_rides=n()
             ,average_duration = mean(ride_length)) %>% 	
  arrange(usertype, weekday)	
  

```

```{r analyzing data}

tripdata_df_v2 %>% 
  mutate(weekday = wday(start_time, label = TRUE)) %>% 
  group_by(usertype, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(usertype, weekday)  %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = usertype)) +
  geom_col(position = "dodge")


```




```{r}

tripdata_df_v2 %\>% mutate(weekday = wday(start_time, label = TRUE))
%\>% group_by(usertype, weekday) %\>% summarise(number_of_rides = n()
,average_duration = mean(ride_length)) %\>% arrange(usertype, weekday)
%\>% ggplot(aes(x = weekday, y = average_duration, fill = usertype)) +
geom_col(position = "dodge")

```

